A Formal Model of Checked C

### 2023-04-07 sandp23/background.tex
Adding

\section{Background and Motivation}
%\section{Overview and Transcendence}\label{sec:overview}
%This section describes \checkedc{} and new features \systemname{} provides.
Here, we brief \checkedc and the motivation for~\systemname{}.

>>>>

\section{Background and Motivation}
\aravind{I am working on this.}
%\section{Overview and Transcendence}\label{sec:overview}
%This section describes \checkedc{} and new features \systemname{} provides.

\subsection{Memory Safety}
This property ensures that all memory accesses a program makes to valid areas containing the program's data~\cite{payer2019software}.
Languages that guarantee (or do not) this property are called memory-safe (or memory-unsafe) languages.~\eg Java is a memory-safe language, whereas C/C++ is not.
There are two main classes of memory safety violations: spatial and temporal.
Spatial violations occur when a program dereferences a pointer that points outside of the associated object.~\eg buffer overread, underread, array out-of-bounds access, null-ptr dereference, etc.
Temporal violations occur when a program dereferences a pointer that points to an invalid object which was previously object.~\eg use-after-free, double-free, etc.

\subsection{Program Partioning}
\label{subsec:background:programpart}
Program partitioning~\cite{rul2009towards} is a well-known technique for dividing a program into multiple isolated parts or partitions.
There has been considerable work~\cite{tan2017principles, brumley2004privtrans, bittau2008wedge, lind2017glamdring, liu2017ptrsplit} in the area.
Most of these techniques are~\emph{data-centric}~\cite{lind2017glamdring, liu2017ptrsplit}, wherein program data drives the partitioning.
Specifically, given sensitive data in a program, all functions accessing sensitive data will be considered part of the trusted partition, and all other functions will be in the untrusted partition.
For instance, in the following code~\code{key} is marked as sensitive and functions that access it,~\ie{}~\code{initkey} and~\code{encrypt} are considered to be part of trusted parition and the other function,~\ie{}~\code{main} will be in the untrusted partition.
\begin{minted}[xleftmargin=25pt, mathescape, linenos, escapeinside=||, fontsize=\footnotesize{}]{c}
char __attribute__((annotate("sensitive"))) *key;

void initkey(int sz) {
 ...
 for (i=0; i<sz; i++) key[i]= ...;
}

void encrypt(char *ptxt, int sz) {
 ..
 for (i=0; i<sz; i++)
  ctxt[i]=ptxt[i] ^ key[i];
 ..
}

int main() {
 ..
 initkey(strlen(txt));
 encrypt(txt, strlen(txt));
 ... 
}
\end{minted}

As we can see, partitioning is achieved easily in the existing techniques by identifying which functions access sensitive data.
These partitions can be separated in various ways,~\eg{} as different processes communicating through~\ac{RPC}~\cite{nelson1981remote}, parent and child processes communicating through shared memory, etc.

The main novelty of existing techniques is using various program analysis and dynamic instrumentation methods to facilitate transparent communication between functions in trusted and untrusted partitions.
Because of their focus on inter-partition interactions, the existing techniques are specialized for the target isolation mechanism. For instance,~\ptrsplit{}~\cite{liu2017ptrsplit} technique assumes that partitions are hosted as separate processes and data is exchanged through marshaling,~\eg{} sending data pointed by~\code{ptxt} and~\code{sz} during calls to~\code{encrypt} (\ie line 18 in the above example) and synchronizing the data back to~\code{ctxt} on return.
Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.
Similarly,~\glamdring~\cite{lind2017glamdring} is specialized for partitioning using SGX-enclaves.

However, these isolation mechanisms may not be available on all systems.
Furthermore, the existing partioning techniques focus on sensitive data use-case -- where the main goal is to protect the integrity and confidentiality of the sensitive data.
Such isolation mechanisms might be unnecessary or too expensive for security use cases.
For instance, consider the case of isolating functions (\eg input parsing) that process tainted data. 
Here, we do not care about the integrity or confidentiality of the tainted data but rather want to restrict the effect of these functions on the overall program.

Consequently, developers might want to use lightweight isolation mechanisms with low-performance overhead and limited guarantees (\eg isolation against spatial safety issues).
The existing techniques do not support such flexible use cases because of their tight coupling with the isolation mechanism.

% SFI: mechanisms and its use cases.

% Checked C

% Motivation: vulnerabilities.
Here, we brief \checkedc and the motivation for~\systemname{}.

### 2023-04-07 sandp23/introduction.tex
Adding

One of the main reasons for this is combining partitioning with~\acf{SFI}.
Specifically, in addition to partitioning, existing techniques also provide~\ac{SFI} such that code in the untrusted region cannot affect code in the trusted region,~\eg{} by executing untrusted partition in a separate process~\cite{liu2017ptrsplit} or executing in a Trusted Execution environment~\cite{lind2017glamdring}.
However, this extreme partitioning is not always necessary and feasible in certain resource constrained environments.
For instance, we want to isolate a string processing function only from spatial violations,~\ie any spatial violations (\eg buffer overread) in that function should not affect the program.

But existing partitioning techniques are specialized for a specific~\ac{SFI}~\cite{wahbe1993efficient} technique.
For instance,~\ptrsplit{}~\cite{liu2017ptrsplit} technique assumes that partitions are hosted as separate processes and data is exchanged through marshaling.
Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.
In summary, existing techniques combine policy (\ie{}~\emph{what} to partition) with mechanism (\ie{}~\emph{how} to partition).
We argue that existing program partitioning mechanisms are restrictive and inflexible for different use cases (\eg isolation from spatial safety vulnerabilities). We need a technique that enables developers to flexibility partition programs and enforce isolation -- balancing safety guarantees and performance overhead.

>>>>

One of the main reasons for this is combining partitioning with~\acf{SFI}~\cite{wahbe1993efficient}.
Specifically, in addition to partitioning, existing techniques also provide~\ac{SFI} such that code in the untrusted region cannot affect code in the trusted region,~\eg{} by executing untrusted partition in a separate process~\cite{liu2017ptrsplit} or executing in a Trusted Execution environment~\cite{lind2017glamdring}.
However, this extreme partitioning is not always necessary and feasible in certain resource constrained environments.
For instance, we want to isolate a string processing function only from spatial violations,~\ie any spatial violations (\eg buffer overread) in that function should not affect the program.
%
But existing partitioning techniques are specialized for a specific~\ac{SFI} technique (\sect{subsec:background:programpart}).
%For instance,~\ptrsplit{}~\cite{liu2017ptrsplit} technique assumes that partitions are hosted as separate processes and data is exchanged through marshaling.
%Any other isolation mechanism or data exchange (\eg shared memory) cannot be supported as the partitioning technique is specialized for marshaling by using~\ac{PDG} and tracking bounds information.

In other words, existing techniques combine policy (\ie{}~\emph{what} to partition) with mechanism (\ie{}~\emph{how} to partition).
We argue that existing program partitioning mechanisms are restrictive and inflexible for different use cases (\eg isolation from spatial safety vulnerabilities). We need a technique that enables developers to flexibility partition programs and enforce isolation -- balancing safety guarantees and performance overhead.
